---
title: 'Viewership Report: Suits'
author: "Elvis Udabor"
date: "2026-04-16"
format:
  html:
    toc: true
    theme: flatly
execute:
  echo: false
---

```{r}
#| message: false
#| warning: false

library(tidyverse)
library(rvest)
library(janitor)
library(purrr)
library(lubridate)
library(knitr)

url <- "https://en.wikipedia.org/wiki/List_of_Suits_episodes"
page <- read_html(url)

tables <- page |>
  html_elements("table.wikiepisodetable") |>
  html_table()

# build one combined episode dataset from all episode tables
episodes <- map_dfr(seq_along(tables), function(i) {
  tab <- tables[[i]] |>
    clean_names()
  
  # keep only tables that contain a viewers column
  if (!any(str_detect(names(tab), "viewers"))) {
    return(NULL)
  }
  
  tab$season <- i
  tab
})
```



## About the show

Suits is an American television drama series created by Aaron Korsh. It premiered on USA Network in 2011 and follows Mike Ross, who begins working as a law associate for Harvey Specter despite never attending law school. The series ran for nine seasons and 134 episodes.

## Data

```{r}
episodes_clean <- episodes |>
  clean_names()

# identify likely columns
viewer_col <- names(episodes_clean)[str_detect(names(episodes_clean), "viewers")][1]
date_col   <- names(episodes_clean)[str_detect(names(episodes_clean), "original_air_date|original_release_date")][1]
season_ep_col <- names(episodes_clean)[str_detect(names(episodes_clean), "^no_in_season$|season_no|episode_no")][1]
overall_ep_col <- names(episodes_clean)[str_detect(names(episodes_clean), "^no_overall$|overall")][1]

# rename only the columns that actually exist
if (!is.na(viewer_col)) {
  episodes_clean <- episodes_clean |> rename(viewers_millions = all_of(viewer_col))
}

if (!is.na(date_col)) {
  episodes_clean <- episodes_clean |> rename(release_date = all_of(date_col))
}

if (!is.na(season_ep_col)) {
  episodes_clean <- episodes_clean |> rename(episode_in_season = all_of(season_ep_col))
}

if (!is.na(overall_ep_col)) {
  episodes_clean <- episodes_clean |> rename(episode_overall = all_of(overall_ep_col))
}

# clean required columns
episodes_clean <- episodes_clean |>
  mutate(
    viewers_millions = str_extract(as.character(viewers_millions), "\\d+\\.?\\d*") |> as.numeric(),
    release_date_raw = str_remove_all(as.character(release_date), "\\[.*?\\]") |> str_squish(),
    release_date = case_when(
      str_detect(release_date_raw, "\\d{4}-\\d{2}-\\d{2}") ~
        str_extract(release_date_raw, "\\d{4}-\\d{2}-\\d{2}") |> ymd(),
      TRUE ~ parse_date_time(release_date_raw, orders = c("mdy", "dmy"))
    )
  )

# clean optional episode-number columns only if present
if ("episode_in_season" %in% names(episodes_clean)) {
  episodes_clean <- episodes_clean |>
    mutate(
      episode_in_season = str_extract(as.character(episode_in_season), "\\d+") |> as.integer()
    )
}

if ("episode_overall" %in% names(episodes_clean)) {
  episodes_clean <- episodes_clean |>
    mutate(
      episode_overall = str_extract(as.character(episode_overall), "\\d+") |> as.integer()
    )
}

episodes_clean <- episodes_clean |>
  filter(!is.na(viewers_millions), !is.na(release_date)) |>
  arrange(release_date)
```

## Summary statistics

```{r}
summary_stats <- episodes_clean |>
  summarise(
    episodes = n(),
    first_episode_viewers = first(viewers_millions),
    last_episode_viewers = last(viewers_millions),
    average_viewers = mean(viewers_millions),
    median_viewers = median(viewers_millions),
    max_viewers = max(viewers_millions),
    min_viewers = min(viewers_millions)
  )

kable(round(summary_stats, 2), caption = "Basic viewership summary for *Suits*")
```

```{r}
season_stats <- episodes_clean |>
  group_by(season) |>
  summarise(
    avg_viewers = mean(viewers_millions),
    start_viewers = first(viewers_millions),
    end_viewers = last(viewers_millions),
    .groups = "drop"
  )

kable(round(season_stats, 2), caption = "Average viewership by season")
```

### Viewership over time

```{r}
ggplot(episodes_clean, aes(x = release_date, y = viewers_millions)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 1.6) +
  labs(
    title = "Viewership over time for *Suits*",
    x = "Release date",
    y = "U.S. viewers (millions)"
  ) +
  theme_minimal()
```

### Episode-to-episode changes

```{r}
episodes_change <- episodes_clean |>
  arrange(release_date) |>
  mutate(viewer_change = viewers_millions - lag(viewers_millions))

ggplot(episodes_change, aes(x = release_date, y = viewer_change)) +
  geom_hline(yintercept = 0, linewidth = 0.4) +
  geom_col() +
  labs(
    title = "Episode-to-episode change in viewership",
    x = "Release date",
    y = "Change in viewers (millions)"
  ) +
  theme_minimal()
```

### Interpretation

```{r}
first_view <- summary_stats$first_episode_viewers
last_view <- summary_stats$last_episode_viewers
avg_view <- summary_stats$average_viewers
max_view <- summary_stats$max_viewers
min_view <- summary_stats$min_viewers

season_1_avg <- season_stats |> filter(season == 1) |> pull(avg_viewers)
season_5_avg <- season_stats |> filter(season == 5) |> pull(avg_viewers)
season_9_avg <- season_stats |> filter(season == 9) |> pull(avg_viewers)

drop_1_to_5 <- season_5_avg - season_1_avg
drop_1_to_9 <- season_9_avg - season_1_avg
```

The pilot episode of Suits drew r round(first_view, 2) million viewers, while the final observed episode in the scraped data drew r round(last_view, 2) million. Average episode viewership across the series was r round(avg_view, 2) million viewers. The maximum observed audience was r round(max_view, 2) million, while the minimum was r round(min_view, 2) million.

At the season level, average viewership fell from r round(season_1_avg, 2) million in season 1 to r round(season_5_avg, 2) million in season 5, a change of r round(drop_1_to_5, 2) million viewers. By season 9, the average had moved to r round(season_9_avg, 2) million, which is r round(drop_1_to_9, 2) million lower than season 1. Overall, the graphs suggest that the show started with stronger audiences and then generally trended downward over time, although there were still noticeable episode-level fluctuations.

### Source

The report uses the episode list and viewer counts reported on Wikipedia for Suits.
